Race and Threat Priming 
Nick Camp, Takuya Sawaoka, Erica Yoon
========================================================

This document contains signal detection analyses for three experiments: *Weapons Bias* (Study I), *Snake Bias* (Study II), *Door/Window Study* (Study III) and *Control Study* (Study IV).  

```{r, include=FALSE}
rm(list = ls())
## for bootstrapping 95% confidence intervals
theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
ci.low <- function(x,na.rm=T) {
  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
ci.high <- function(x,na.rm=T) {
  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r, include=FALSE}
#libraries
library(RJSONIO)
library(plyr)
library(car)
library(reshape)
library(lme4)
library(lmerTest)
library(ggplot2)
library(bootstrap)
```


#Study I: Replicating Weapons Bias

This study was a direct replication of Payne (2001). Participants saw a Black or White prime for Xms, followed by a gun or a tool. S had to respond within a 500ms response window.

```{r,include=FALSE}
## read in the tab separated values (tsv)
d <- read.delim("/afs/ir.stanford.edu/users/n/c/ncamp/Data/GunsRawData.tsv", stringsAsFactors = FALSE)

to.data.frame <- function(x) {
  do.call(data.frame, as.list(x))
}

## make a long-form data frame containing all the trials
trials1 <- ddply(d, .(workerid), function(worker) {
  
  trials.list.of.vectors <- fromJSON(worker$Answer.data)
  trials.list.of.demographic <- fromJSON(worker$Answer.demographicsData)
  
  ## you can see this "list of vectors" structure by uncommenting the following line:  
  #### print( str(trials.list.of.vectors) )
  
  ## convert every element in trials.list.of.vectors from a vector to a data frame
  ## (do this by "Map"ping the to.data.frame function over trials.list.of.vectors
  trials.list.of.dfs <- Map(to.data.frame, trials.list.of.vectors)
  trials.list.of.demographics <- Map(to.data.frame, trials.list.of.demographic)
  
  ## mush all the of the individual data frames into a single data frame
  trials.df <- do.call(rbind, trials.list.of.dfs) 
  trials.demographics <- do.call(rbind, trials.list.of.demographics) 
  data.frame(
    "workerid" = worker$workerid,
    trials.df,
    trials.demographics
  )
})

# sort on id, race, and tool
weaponData<-trials1[order(trials1$workerid,trials1$race,trials1$tool),]

# rename levels
levels(weaponData$race) <- c("White", "Black")
levels(weaponData$tool) <- c("Tool", "Gun")
```

```{r,include=FALSE}
weaponData = subset(weaponData, Ethnicity != "3")
```

```{r,include=FALSE}
shortformWeapon=cast(weaponData,workerid+Age+Male+Ethnicity+SES+ReligiousAffiliation+ReligiousAttendance+Education+Citizen+Political+Comments~race+tool,value='accuracy', fun.aggregate=mean)
```

###Study I: Computing signal detection stats

For each subject, we calculate bias and sensitivity for BLACK and WHITE targets: 

* Hit is defined as correctly identifying GUN after seeing GUN (accuracy rate for GUN trials)

* False Alarm is defined as incorrectly responding GUN after seeing TOOL (1-accuracy rate for TOOL trials)


```{r}
shortformWeapon$HitRateWhite=shortformWeapon$White_Gun
shortformWeapon$FARateWhite=1-shortformWeapon$White_Tool

shortformWeapon$HitRateBlack=shortformWeapon$Black_Gun
shortformWeapon$FARateBlack=1-shortformWeapon$Black_Tool

#If a S has a HR or FAR of 0 or 1, we can't calculate d prime. This adjusts extreme values by limiting these
#rates to (1/N) and (N-1)/N, where N is the number of trials used to calculate the HR or FAR.

NumTrials=aggregate(tool~workerid+race,data=trials1,length)[1,3];

shortformWeapon$HitRateWhite=recode(shortformWeapon$HitRateWhite,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformWeapon$HitRateBlack=recode(shortformWeapon$HitRateBlack,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")

shortformWeapon$FARateWhite=recode(shortformWeapon$FARateWhite,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformWeapon$FARateBlack=recode(shortformWeapon$FARateBlack,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
```

These functions calculate sensitivity (dPrime), bias (beta), and criterion (c) using the HR and FAR.

```{r}
#D prime and bias functions from Pellier (2002)

#dprime: separation between the signal and the noise distribution, compared against the sd of the noise distribution. Higher dprime means that the signal can more readily be distinguished from the noise.

dprime <- function(hit,fa) {
  qnorm(hit) - qnorm(fa)
}

#beta: bias towards responding that the target is present, independent of whether the target actually is present. Higher bias means greater tendency towards responding 'present'. 

beta <- function(hit,fa) {
  zhr <- qnorm(hit)
  zfar <- qnorm(fa)
  exp(-zhr*zhr/2+zfar*zfar/2)
  
}

#c is defined as the distance between the criterion and the neutral point, where the noise and signal distributions cross over. Negative values of c signify a bias toward responding yes, whereas positive values signify a bias toward the no response (thecriterion lies to the right of the neutral point).

criterion<-function(hit,fa) {
  zhr <- qnorm(hit)
  zfar <- qnorm(fa)
  -(zfar+zhr)/2
}
```

Now we compute sensitivity, bias, and criterion for Black and White primes.
```{r}
shortformWeapon$dPrimeWhite=dprime(shortformWeapon$HitRateWhite,shortformWeapon$FARateWhite)
shortformWeapon$dPrimeBlack=dprime(shortformWeapon$HitRateBlack,shortformWeapon$FARateBlack)

shortformWeapon$biasWhite=beta(shortformWeapon$HitRateWhite,shortformWeapon$FARateWhite)
shortformWeapon$biasBlack=beta(shortformWeapon$HitRateBlack,shortformWeapon$FARateBlack)

shortformWeapon$criterionWhite=criterion(shortformWeapon$HitRateWhite,shortformWeapon$FARateWhite)
shortformWeapon$criterionBlack=criterion(shortformWeapon$HitRateBlack,shortformWeapon$FARateBlack)
```

###Study I: Sensitivity by prime.

```{r, include=FALSE}
dPrimesWeapons=melt(data.frame(subset(shortformWeapon,select=c("workerid","dPrimeWhite","dPrimeBlack")),id.vars="workerid"))
dPrimesWeapons=dPrimesWeapons[,c(1,3,4)]
names(dPrimesWeapons)=c("workerid","race","dPrime")
dPrimesWeapons$race=recode(dPrimesWeapons$race,"'dPrimeWhite'='White';'dPrimeBlack'='Black'")
```

```{r,message=FALSE,echo=FALSE}
ms <- aggregate(dPrime ~ race, dPrimesWeapons, mean)
mss <- aggregate(dPrime ~ race + workerid, dPrimesWeapons, mean)
# for CI's
ms$cih <- aggregate(dPrime ~ race, mss, ci.high)$dPrime
ms$cil <- aggregate(dPrime ~ race, mss, ci.low)$dPrime

s1dPrime <- qplot(race,dPrime,
                 fill=race, stat="identity",
                 geom="bar",
                 data=ms) +
  geom_errorbar(aes(ymin=dPrime-cil,ymax=dPrime+cih,width=.2)) +
  xlab(NULL) + ylab("dPrime") +theme(legend.position="none")+ ggtitle("Sensitivity Responding 'Gun'")
s1dPrime
```

Stats in lmer:
```{r}
weapon.sensitivity <- lmer(dPrime ~ race + (1|workerid), data = dPrimesWeapons);summary(weapon.sensitivity)
```

###Study I: Bias by prime.
```{r,include=FALSE}
biasWeapon=melt(data.frame(subset(shortformWeapon,select=c("workerid","biasWhite","biasBlack")),id.vars="workerid"))
biasWeapon=biasWeapon[,c(1,3,4)]
names(biasWeapon)=c("workerid","race","bias")
biasWeapon$race=recode(biasWeapon$race,"'biasWhite'='White';'biasBlack'='Black'")
```

```{r,message=FALSE,echo=FALSE}
ms <- aggregate(bias ~ race, biasWeapon, mean)
mss <- aggregate(bias ~ race + workerid, biasWeapon, mean)
# for CI's
ms$cih <- aggregate(bias ~ race, mss, ci.high)$bias
ms$cil <- aggregate(bias ~ race, mss, ci.low)$bias

s1bias <- qplot(race,bias,
                 fill=race, stat="identity",
                 geom="bar",
                 data=ms) +
  geom_errorbar(aes(ymin=bias-cil,ymax=bias+cih,width=.2)) +
  xlab(NULL) + ylab("bias") +theme(legend.position="none")+
  ggtitle("Bias Responding 'Gun'")
s1bias
```

Stats in lmer:
```{r}
weapon.bias <- lmer(bias ~ race + (1|workerid), data = biasWeapon);summary(weapon.bias)
```

###Study I: Criterion by prime.

```{r,echo=FALSE,message=FALSE}
criterionWeapon=melt(data.frame(subset(shortformWeapon,select=c("workerid","criterionWhite","criterionBlack")),id.vars="workerid"))
criterionWeapon=criterionWeapon[,c(1,3,4)]
names(criterionWeapon)=c("workerid","race","criterion")
criterionWeapon$race=recode(criterionWeapon$race,"'criterionWhite'='White';'criterionBlack'='Black'")
```

```{r,echo=FALSE,message=FALSE}
ms <- aggregate(criterion ~ race, criterionWeapon, mean)
mss <- aggregate(criterion ~ race + workerid, criterionWeapon, mean)
# for CI's
ms$cih <- aggregate(criterion ~ race, mss, ci.high)$criterion
ms$cil <- aggregate(criterion ~ race, mss, ci.low)$criterion

s1criterion <- qplot(race,criterion,
               fill=race, stat="identity",
               geom="bar",
               data=ms) +
  geom_errorbar(aes(ymin=criterion-cil,ymax=criterion+cih,width=.2)) +
  xlab(NULL) + ylab("criterion") +theme(legend.position="none")+
  ggtitle("Criterion Responding 'Gun'")
s1criterion
```

```{r}
weapon.criterion <- lmer(criterion ~ race + (1|workerid), data = criterionWeapon);summary(weapon.criterion)

#Question: not sure why the workerid variance=0 in the Random Effects. Shouldn't the model be able to fit a participant level intercept across Black and White prime conditions?
```

Summary: When targets were primed with Black faces, participants were marginally LESS accurate in identifying weapons. There was no difference in bias, but a difference in criterion for responding "gun": participants set a lower criterion after seeing Black faces.

#Study II: Race primes Snake identification

This study was the same design as Study I, only with Snakes substituted for Guns, and Ropes substituted for Tools.

Data setup is hidden, as it is the same as in Study I.
```{r,message=FALSE,echo=FALSE}
## read in the tab separated values (tsv)
d <- read.delim("/afs/ir.stanford.edu/users/n/c/ncamp/Data/SnakesRawData.tsv", stringsAsFactors = FALSE)

to.data.frame <- function(x) {
  do.call(data.frame, as.list(x))
}

## make a long-form data frame containing all the trials
trials1 <- ddply(d, .(workerid), function(worker) {
  
  trials.list.of.vectors <- fromJSON(worker$Answer.data)
  trials.list.of.demographic <- fromJSON(worker$Answer.demographicsData)
  
  ## you can see this "list of vectors" structure by uncommenting the following line:  
  #### print( str(trials.list.of.vectors) )
  
  ## convert every element in trials.list.of.vectors from a vector to a data frame
  ## (do this by "Map"ping the to.data.frame function over trials.list.of.vectors
  trials.list.of.dfs <- Map(to.data.frame, trials.list.of.vectors)
  trials.list.of.demographics <- Map(to.data.frame, trials.list.of.demographic)
  
  ## mush all the of the individual data frames into a single data frame
  trials.df <- do.call(rbind, trials.list.of.dfs) 
  trials.demographics <- do.call(rbind, trials.list.of.demographics) 
  data.frame(
    "workerid" = worker$workerid,
    trials.df,
    trials.demographics
  )
})

# sort on id, race, and snake
snakeData<-trials1[order(trials1$workerid,trials1$race,trials1$snake),]

# rename levels
levels(snakeData$race) <- c("White", "Black")
levels(snakeData$snake) <- c("Rope", "Snake")
```

```{r, message=FALSE,echo=FALSE}
snakeData = subset(snakeData, Ethnicity != "3")
```

```{r, message=FALSE,echo=FALSE}
shortformSnake=cast(snakeData,workerid+Age+Male+Ethnicity+SES+ReligiousAffiliation+ReligiousAttendance+Education+Citizen+Political+Comments~race+snake,value='accuracy', fun.aggregate=mean)
```

###Study II: Computing signal detection stats

For each subject, we calculate bias and sensitivity for BLACK and WHITE targets: 

* Hit is defined as correctly identifying SNAKE after seeing SNAKE (accuracy rate for SNAKE trials)

* False Alarm is defined as incorrectly responding SNAKE after seeing ROPE (1-accuracy rate for ROPE trials)

*Omitted syntax below, since it's so similar to Study I

```{r,echo=FALSE,message=FALSE}
shortformSnake$HitRateWhite=shortformSnake$White_Snake
shortformSnake$FARateWhite=1-shortformSnake$White_Rope

shortformSnake$HitRateBlack=shortformSnake$Black_Snake
shortformSnake$FARateBlack=1-shortformSnake$Black_Rope

#If a S has a HR or FAR of 0 or 1, we can't calculate d prime. This adjusts extreme values by limiting these
#rates to (1/N) and (N-1)/N, where N is the number of trials used to calculate the HR or FAR.

NumTrials=aggregate(snake~workerid+race,data=trials1,length)[1,3];

shortformSnake$HitRateWhite=recode(shortformSnake$HitRateWhite,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformSnake$HitRateBlack=recode(shortformSnake$HitRateBlack,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")

shortformSnake$FARateWhite=recode(shortformSnake$FARateWhite,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformSnake$FARateBlack=recode(shortformSnake$FARateBlack,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
```


Compute sensitivity, bias, and criterion for Black and White primes...
```{r,message=FALSE,echo=FALSE}
shortformSnake$dPrimeWhite=dprime(shortformSnake$HitRateWhite,shortformSnake$FARateWhite)
shortformSnake$dPrimeBlack=dprime(shortformSnake$HitRateBlack,shortformSnake$FARateBlack)

shortformSnake$biasWhite=beta(shortformSnake$HitRateWhite,shortformSnake$FARateWhite)
shortformSnake$biasBlack=beta(shortformSnake$HitRateBlack,shortformSnake$FARateBlack)

shortformSnake$criterionWhite=criterion(shortformSnake$HitRateWhite,shortformSnake$FARateWhite)
shortformSnake$criterionBlack=criterion(shortformSnake$HitRateBlack,shortformSnake$FARateBlack)
```

###Study II: Sensitivity by prime.

```{r,message=FALSE,echo=FALSE}
dPrimesSnakes=melt(data.frame(subset(shortformSnake,select=c("workerid","dPrimeWhite","dPrimeBlack")),id.vars="workerid"))
dPrimesSnakes=dPrimesSnakes[,c(1,3,4)]
names(dPrimesSnakes)=c("workerid","race","dPrime")
dPrimesSnakes$race=recode(dPrimesSnakes$race,"'dPrimeWhite'='White';'dPrimeBlack'='Black'")
```

```{r, echo=FALSE}
ms <- aggregate(dPrime ~ race, dPrimesSnakes, mean)
mss <- aggregate(dPrime ~ race + workerid, dPrimesSnakes, mean)
# for CI's
ms$cih <- aggregate(dPrime ~ race, mss, ci.high)$dPrime
ms$cil <- aggregate(dPrime ~ race, mss, ci.low)$dPrime

s2dPrime <- qplot(race,dPrime,
                 fill=race, stat="identity",
                 geom="bar",
                 data=ms) +
  geom_errorbar(aes(ymin=dPrime-cil,ymax=dPrime+cih,width=.2)) +
  xlab(NULL) + ylab("dPrime") +theme(legend.position="none")+coord_cartesian(ylim = c(-.75, .75))+ggtitle("Sensitivity Responding 'Snake'")
s2dPrime
```

Stats in lmer:
```{r}
snake.sensitivity <- lmer(dPrime ~ race + (1|workerid), data = dPrimesSnakes);summary(snake.sensitivity)
```

###Study II: Bias by prime.
```{r,echo=FALSE,message=FALSE}
biasSnake=melt(data.frame(subset(shortformSnake,select=c("workerid","biasWhite","biasBlack")),id.vars="workerid"))
biasSnake=biasSnake[,c(1,3,4)]
names(biasSnake)=c("workerid","race","bias")
biasSnake$race=recode(biasSnake$race,"'biasWhite'='White';'biasBlack'='Black'")
```

```{r,echo=FALSE}
ms <- aggregate(bias ~ race, biasSnake, mean)
mss <- aggregate(bias ~ race + workerid, biasSnake, mean)
# for CI's
ms$cih <- aggregate(bias ~ race, mss, ci.high)$bias
ms$cil <- aggregate(bias ~ race, mss, ci.low)$bias

s2bias <- qplot(race,bias,
                 fill=race, stat="identity",
                 geom="bar",
                 data=ms) +
  geom_errorbar(aes(ymin=bias-cil,ymax=bias+cih,width=.2)) +
  xlab(NULL) + ylab("bias") +theme(legend.position="none")+
  ggtitle("Bias for Responding 'Snake'")
s2bias
```

Stats in lmer:
```{r}
snake.bias <- lmer(bias ~ race + (1|workerid), data = biasSnake);summary(snake.bias)
```

###Study II: Criterion by prime.

```{r,message=FALSE,echo=FALSE}
criterionSnake=melt(data.frame(subset(shortformSnake,select=c("workerid","criterionWhite","criterionBlack")),id.vars="workerid"))
criterionSnake=criterionSnake[,c(1,3,4)]
names(criterionSnake)=c("workerid","race","criterion")
criterionSnake$race=recode(criterionSnake$race,"'criterionWhite'='White';'criterionBlack'='Black'")
```

```{r,echo=FALSE,message=FALSE}

ms <- aggregate(criterion ~ race, criterionSnake, mean)
mss <- aggregate(criterion ~ race + workerid, criterionSnake, mean)
# for CI's
ms$cih <- aggregate(criterion ~ race, mss, ci.high)$criterion
ms$cil <- aggregate(criterion ~ race, mss, ci.low)$criterion

s2criterion <- qplot(race,criterion,
               fill=race, stat="identity",
               geom="bar",
               data=ms) +
  geom_errorbar(aes(ymin=criterion-cil,ymax=criterion+cih,width=.2)) +
  xlab(NULL) + ylab("criterion") +theme(legend.position="none")+
  ggtitle("Criterion Responding 'Snake' ")
s2criterion
```

Stats with lmer:
```{r}
snake.criterion <- lmer(criterion ~ race + (1|workerid), data = criterionSnake);summary(snake.criterion)
```

##Summary: When targets were primed with Black faces, participants set a lower criterion for responding "snake". There were no differences in bias or sensitivity.

##Door/Window Study
```{r,message=FALSE,echo=FALSE}
## read in the tab separated values (tsv)
d <- read.delim("/afs/ir.stanford.edu/users/n/c/ncamp/Data/DoorsRawData.tsv", stringsAsFactors = FALSE)

to.data.frame <- function(x) {
  do.call(data.frame, as.list(x))
}

## make a long-form data frame containing all the trials
trials1 <- ddply(d, .(workerid), function(worker) {
  
  trials.list.of.vectors <- fromJSON(worker$Answer.data)
  trials.list.of.demographic <- fromJSON(worker$Answer.demographicsData)
  
  ## you can see this "list of vectors" structure by uncommenting the following line:  
  #### print( str(trials.list.of.vectors) )
  
  ## convert every element in trials.list.of.vectors from a vector to a data frame
  ## (do this by "Map"ping the to.data.frame function over trials.list.of.vectors
  trials.list.of.dfs <- Map(to.data.frame, trials.list.of.vectors)
  trials.list.of.demographics <- Map(to.data.frame, trials.list.of.demographic)
  
  ## mush all the of the individual data frames into a single data frame
  trials.df <- do.call(rbind, trials.list.of.dfs) 
  trials.demographics <- do.call(rbind, trials.list.of.demographics) 
  data.frame(
    "workerid" = worker$workerid,
    trials.df,
    trials.demographics
  )
})

# sort on id, race, and door
doorData<-trials1[order(trials1$workerid,trials1$race,trials1$door),]

# rename levels
levels(doorData$race) <- c("White", "Black")
levels(doorData$door) <- c("window", "door")
```

```{r, message=FALSE,echo=FALSE}
doorData = subset(doorData, Ethnicity != "3")
```

```{r, message=FALSE,echo=FALSE}
shortformdoor=cast(doorData,workerid+Age+Male+Ethnicity+SES+ReligiousAffiliation+ReligiousAttendance+Education+Citizen+Political+Comments~race+door,value='accuracy', fun.aggregate=mean)
```

###Study III: Computing signal detection stats

For each subject, we calculate bias and sensitivity for BLACK and WHITE targets: 

* Hit is defined as correctly identifying door after seeing door (accuracy rate for door trials)

* False Alarm is defined as incorrectly responding door after seeing window (1-accuracy rate for window trials)

*Omitted syntax below, since it's so similar to Study I

```{r,echo=FALSE,message=FALSE}
shortformdoor$HitRateWhite=shortformdoor$White_door
shortformdoor$FARateWhite=1-shortformdoor$White_window

shortformdoor$HitRateBlack=shortformdoor$Black_door
shortformdoor$FARateBlack=1-shortformdoor$Black_window

#If a S has a HR or FAR of 0 or 1, we can't calculate d prime. This adjusts extreme values by limiting these
#rates to (1/N) and (N-1)/N, where N is the number of trials used to calculate the HR or FAR.

NumTrials=aggregate(door~workerid+race,data=trials1,length)[1,3];

shortformdoor$HitRateWhite=recode(shortformdoor$HitRateWhite,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformdoor$HitRateBlack=recode(shortformdoor$HitRateBlack,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")

shortformdoor$FARateWhite=recode(shortformdoor$FARateWhite,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformdoor$FARateBlack=recode(shortformdoor$FARateBlack,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
```


Compute sensitivity, bias, and criterion for Black and White primes...
```{r,message=FALSE,echo=FALSE}
shortformdoor$dPrimeWhite=dprime(shortformdoor$HitRateWhite,shortformdoor$FARateWhite)
shortformdoor$dPrimeBlack=dprime(shortformdoor$HitRateBlack,shortformdoor$FARateBlack)

shortformdoor$biasWhite=beta(shortformdoor$HitRateWhite,shortformdoor$FARateWhite)
shortformdoor$biasBlack=beta(shortformdoor$HitRateBlack,shortformdoor$FARateBlack)

shortformdoor$criterionWhite=criterion(shortformdoor$HitRateWhite,shortformdoor$FARateWhite)
shortformdoor$criterionBlack=criterion(shortformdoor$HitRateBlack,shortformdoor$FARateBlack)
```

###Study III: Sensitivity by prime.

```{r,message=FALSE,echo=FALSE}
dPrimesdoors=melt(data.frame(subset(shortformdoor,select=c("workerid","dPrimeWhite","dPrimeBlack")),id.vars="workerid"))
dPrimesdoors=dPrimesdoors[,c(1,3,4)]
names(dPrimesdoors)=c("workerid","race","dPrime")
dPrimesdoors$race=recode(dPrimesdoors$race,"'dPrimeWhite'='White';'dPrimeBlack'='Black'")
```

```{r, echo=FALSE}
ms <- aggregate(dPrime ~ race, dPrimesdoors, mean)
mss <- aggregate(dPrime ~ race + workerid, dPrimesdoors, mean)
# for CI's
ms$cih <- aggregate(dPrime ~ race, mss, ci.high)$dPrime
ms$cil <- aggregate(dPrime ~ race, mss, ci.low)$dPrime

s3dPrime <- qplot(race,dPrime,
                 fill=race, stat="identity",
                 geom="bar",
                 data=ms) +
  geom_errorbar(aes(ymin=dPrime-cil,ymax=dPrime+cih,width=.2)) +
  xlab(NULL) + ylab("dPrime") +theme(legend.position="none")+
  ggtitle("Sensitivity Responding 'door'")
s3dPrime
```

Stats in lmer:
```{r}
door.sensitivity <- lmer(dPrime ~ race + (1|workerid), data = dPrimesdoors);summary(door.sensitivity)
```

###Study III: Bias by prime.
```{r,echo=FALSE,message=FALSE}
biasdoor=melt(data.frame(subset(shortformdoor,select=c("workerid","biasWhite","biasBlack")),id.vars="workerid"))
biasdoor=biasdoor[,c(1,3,4)]
names(biasdoor)=c("workerid","race","bias")
biasdoor$race=recode(biasdoor$race,"'biasWhite'='White';'biasBlack'='Black'")
```

```{r,echo=FALSE}
ms <- aggregate(bias ~ race, biasdoor, mean)
mss <- aggregate(bias ~ race + workerid, biasdoor, mean)
# for CI's
ms$cih <- aggregate(bias ~ race, mss, ci.high)$bias
ms$cil <- aggregate(bias ~ race, mss, ci.low)$bias

s3bias <- qplot(race,bias,
                 fill=race, stat="identity",
                 geom="bar",
                 data=ms) +
  geom_errorbar(aes(ymin=bias-cil,ymax=bias+cih,width=.2)) +
  xlab(NULL) + ylab("bias") +theme(legend.position="none")+
  ggtitle("Bias for Responding 'door'")
s3bias
```

Stats in lmer:
```{r}
door.bias <- lmer(bias ~ race + (1|workerid), data = biasdoor);summary(door.bias)
```

###Study III: Criterion by prime.

```{r,message=FALSE,echo=FALSE}
criteriondoor=melt(data.frame(subset(shortformdoor,select=c("workerid","criterionWhite","criterionBlack")),id.vars="workerid"))
criteriondoor=criteriondoor[,c(1,3,4)]
names(criteriondoor)=c("workerid","race","criterion")
criteriondoor$race=recode(criteriondoor$race,"'criterionWhite'='White';'criterionBlack'='Black'")
```

```{r,echo=FALSE,message=FALSE}

ms <- aggregate(criterion ~ race, criteriondoor, mean)
mss <- aggregate(criterion ~ race + workerid, criteriondoor, mean)
# for CI's
ms$cih <- aggregate(criterion ~ race, mss, ci.high)$criterion
ms$cil <- aggregate(criterion ~ race, mss, ci.low)$criterion

s3criterion <- qplot(race,criterion,
               fill=race, stat="identity",
               geom="bar",
               data=ms) +
  geom_errorbar(aes(ymin=criterion-cil,ymax=criterion+cih,width=.2)) +
  xlab(NULL) + ylab("criterion") +theme(legend.position="none")+
  ggtitle("Criterion Responding 'door' ")
s3criterion
```

Stats with lmer:
```{r}
door.criterion <- lmer(criterion ~ race + (1|workerid), data = criteriondoor);summary(door.criterion)
```

##Summary: There is no race priming effect for doors (versus windows) for sensitivity, bias, or criterion. 


##Control Study: No Race Prime

In the control study, S (N=40) had 80 trials where gun or tool images were preceded by an image of static where the face had been in Studies I and II.

To help interpret the control condition, I'm graphing the results next to the Study I results.


```{r,echo=FALSE,message=FALSE}
## read in the tab separated values (tsv)
d <- read.delim("/afs/ir.stanford.edu/users/n/c/ncamp/Data/GunsControlRawData.tsv", stringsAsFactors = FALSE)

to.data.frame <- function(x) {
  do.call(data.frame, as.list(x))
}

## make a long-form data frame containing all the trials
trials1 <- ddply(d, .(workerid), function(worker) {
  
  trials.list.of.vectors <- fromJSON(worker$Answer.data)
  trials.list.of.demographic <- fromJSON(worker$Answer.demographicsData)
  
  ## you can see this "list of vectors" structure by uncommenting the following line:  
  #### print( str(trials.list.of.vectors) )
  
  ## convert every element in trials.list.of.vectors from a vector to a data frame
  ## (do this by "Map"ping the to.data.frame function over trials.list.of.vectors
  trials.list.of.dfs <- Map(to.data.frame, trials.list.of.vectors)
  trials.list.of.demographics <- Map(to.data.frame, trials.list.of.demographic)
  
  ## mush all the of the individual data frames into a single data frame
  trials.df <- do.call(rbind, trials.list.of.dfs) 
  trials.demographics <- do.call(rbind, trials.list.of.demographics) 
  data.frame(
    "workerid" = worker$workerid,
    trials.df,
    trials.demographics
  )
})

# sort on id, race, and tool
weaponControlData<-trials1[order(trials1$workerid,trials1$tool),]

# rename levels
levels(weaponControlData$tool) <- c("Tool", "Gun")


shortformweaponControl=cast(weaponControlData,workerid+Age+Male+Ethnicity+SES+ReligiousAffiliation+ReligiousAttendance+Education+Citizen+Political+Comments~tool,value='accuracy', fun.aggregate=mean)

shortformweaponControl$HitRate=shortformweaponControl$Gun
shortformweaponControl$FARate=1-shortformweaponControl$Tool

#If a S has a HR or FAR of 0 or 1, we can't calculate d prime. This adjusts extreme values by limiting these
#rates to (1/N) and (N-1)/N, where N is the number of trials used to calculate the HR or FAR.

NumTrials=aggregate(tool~workerid,data=trials1,length)[1,2];

shortformweaponControl$HitRate=recode(shortformweaponControl$HitRate,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")
shortformweaponControl$FARate=recode(shortformweaponControl$FARate,"0=(1/NumTrials);1=((NumTrials-1)/NumTrials)")

shortformweaponControl$dPrime=dprime(shortformweaponControl$HitRate,shortformweaponControl$FARate)
shortformweaponControl$bias=beta(shortformweaponControl$HitRate,shortformweaponControl$FARate)
shortformweaponControl$criterion=criterion(shortformweaponControl$HitRate,shortformweaponControl$FARate)
```


###Control: Sensitivity.
```{r,echo=FALSE,message=FALSE}

dPrimeWeapon=melt(data.frame(subset(shortformWeapon,select=c("workerid","dPrimeWhite","dPrimeBlack")),id.vars="workerid"))
dPrimeWeapon=dPrimeWeapon[,c(1,3,4)]
names(dPrimeWeapon)=c("workerid","race","dPrime")
dPrimeWeapon$race=recode(dPrimeWeapon$race,"'dPrimeWhite'='White (Study 1)';'dPrimeBlack'='Black (Study 1)'")

ms <- aggregate(dPrime ~ race, dPrimeWeapon, mean)
mss <- aggregate(dPrime ~ race + workerid, dPrimeWeapon, mean)
# for CI's
ms$cih <- aggregate(dPrime ~ race, mss, ci.high)$dPrime
ms$cil <- aggregate(dPrime ~ race, mss, ci.low)$dPrime

#Just going to shove the control condition in here...

dPrimeWeaponControl=melt(data.frame(subset(shortformweaponControl,select=c("workerid","dPrime")),id.vars="workerid"))
dPrimeWeaponControl=dPrimeWeaponControl[,c(1,3,4)]
names(dPrimeWeaponControl)=c("workerid","race","dPrime")
dPrimeWeaponControl$race=recode(dPrimeWeaponControl$race,"'dPrime'='Control'")

ms2 <- aggregate(dPrime ~ race, dPrimeWeaponControl, mean)
mss2 <- aggregate(dPrime ~ race + workerid, dPrimeWeaponControl, mean)
# for CI's
ms2$cih <- aggregate(dPrime ~ race, mss2, ci.high)$dPrime
ms2$cil <- aggregate(dPrime ~ race, mss2, ci.low)$dPrime
ms=rbind(ms,ms2)
cdPrime <- qplot(race,dPrime,
               fill=race, stat="identity",
               geom="bar",
               data=ms) +
  geom_errorbar(aes(ymin=dPrime-cil,ymax=dPrime+cih,width=.2)) +
  xlab(NULL) + ylab("dPrime") + geom_bar(data=ms)+theme(legend.position="none")+
  ggtitle("dPrime Black/White/Control")
cdPrime
```

###Control: Bias.
```{r,echo=FALSE,message=FALSE}

biasWeapon=melt(data.frame(subset(shortformWeapon,select=c("workerid","biasWhite","biasBlack")),id.vars="workerid"))
biasWeapon=biasWeapon[,c(1,3,4)]
names(biasWeapon)=c("workerid","race","bias")
biasWeapon$race=recode(biasWeapon$race,"'biasWhite'='White (Study 1)';'biasBlack'='Black (Study 1)'")

ms <- aggregate(bias ~ race, biasWeapon, mean)
mss <- aggregate(bias ~ race + workerid, biasWeapon, mean)
# for CI's
ms$cih <- aggregate(bias ~ race, mss, ci.high)$bias
ms$cil <- aggregate(bias ~ race, mss, ci.low)$bias

#Just going to shove the control condition in here...

biasWeaponControl=melt(data.frame(subset(shortformweaponControl,select=c("workerid","bias")),id.vars="workerid"))
biasWeaponControl=biasWeaponControl[,c(1,3,4)]
names(biasWeaponControl)=c("workerid","race","bias")
biasWeaponControl$race=recode(biasWeaponControl$race,"'bias'='Control'")

ms2 <- aggregate(bias ~ race, biasWeaponControl, mean)
mss2 <- aggregate(bias ~ race + workerid, biasWeaponControl, mean)
# for CI's
ms2$cih <- aggregate(bias ~ race, mss2, ci.high)$bias
ms2$cil <- aggregate(bias ~ race, mss2, ci.low)$bias
ms=rbind(ms,ms2)
cbias <- qplot(race,bias,
               fill=race, stat="identity",
               geom="bar",
               data=ms) +
  geom_errorbar(aes(ymin=bias-cil,ymax=bias+cih,width=.2)) +
  xlab(NULL) + ylab("bias") + geom_bar(data=ms)+theme(legend.position="none")+
  ggtitle("bias Black/White/Control")
cbias
```

###Control: Criterion.

```{r,echo=FALSE,message=FALSE}

criterionWeapon=melt(data.frame(subset(shortformWeapon,select=c("workerid","criterionWhite","criterionBlack")),id.vars="workerid"))
criterionWeapon=criterionWeapon[,c(1,3,4)]
names(criterionWeapon)=c("workerid","race","criterion")
criterionWeapon$race=recode(criterionWeapon$race,"'criterionWhite'='White (Study 1)';'criterionBlack'='Black (Study 1)'")

ms <- aggregate(criterion ~ race, criterionWeapon, mean)
mss <- aggregate(criterion ~ race + workerid, criterionWeapon, mean)
# for CI's
ms$cih <- aggregate(criterion ~ race, mss, ci.high)$criterion
ms$cil <- aggregate(criterion ~ race, mss, ci.low)$criterion

#Just going to shove the control condition in here...

criterionWeaponControl=melt(data.frame(subset(shortformweaponControl,select=c("workerid","criterion")),id.vars="workerid"))
criterionWeaponControl=criterionWeaponControl[,c(1,3,4)]
names(criterionWeaponControl)=c("workerid","race","criterion")
criterionWeaponControl$race=recode(criterionWeaponControl$race,"'criterion'='Control'")

ms2 <- aggregate(criterion ~ race, criterionWeaponControl, mean)
mss2 <- aggregate(criterion ~ race + workerid, criterionWeaponControl, mean)
# for CI's
ms2$cih <- aggregate(criterion ~ race, mss2, ci.high)$criterion
ms2$cil <- aggregate(criterion ~ race, mss2, ci.low)$criterion
ms=rbind(ms,ms2)
ccriterion <- qplot(race,criterion,
               fill=race, stat="identity",
               geom="bar",
               data=ms) +
  geom_errorbar(aes(ymin=criterion-cil,ymax=criterion+cih,width=.2)) +
  xlab(NULL) + ylab("criterion") + geom_bar(data=ms)+theme(legend.position="none")+
  ggtitle("Criterion Black/White/Control")
ccriterion
```

##Summary: When objects were primed with static images rather than faces, participants set a criterion that is equal to the the criterion for Black faces in Study I. 

##However, participants were also more sensitive in the control study than they were in Study I. 


#Summary Plots across all studies:

###Sensitivity
```{r,echo=FALSE}
multiplot(cdPrime+coord_cartesian(ylim = c(-.4, .75))+ggtitle("Guns"),s2dPrime+coord_cartesian(ylim = c(-.4, .75))+ggtitle("Snakes"),s3dPrime+coord_cartesian(ylim = c(-.4, .75))+ggtitle("Doors"),cols=3)
```

###Bias
```{r,echo=FALSE}
multiplot(cbias+coord_cartesian(ylim = c(0, 2))+ggtitle("Guns"),s2bias+coord_cartesian(ylim = c(0, 2))+ggtitle("Snakes"),s3bias+coord_cartesian(ylim = c(0, 2))+ggtitle("Doors"),cols=3)
```

###Criterion
```{r,echo=FALSE}
multiplot(ccriterion+coord_cartesian(ylim = c(-.4, .2))+ggtitle("Guns"),s2criterion+coord_cartesian(ylim = c(-.4, .2))+ggtitle("Snakes"),s3criterion+coord_cartesian(ylim = c(-.4, .2))+ggtitle("Doors"),cols=3)
```